---
title: "bacteria_processing_3"
author: "Ilya"
date: "6/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####install and load required packages
```{r packages, echo=FALSE}
source("packages_bacteria1.R")
devtools::install_github("guiastrennec/ggplus")
library(ggplus)
#https://github.com/guiastrennec/ggplus
```


###1. Get bacteria-caused dx in GIDEON   
###2. Match dx to pathogen spp. names
###3. Match spp. names (from GIDEON zdx and GMPD) to traits (in GMPD & other)
###4. Compile master list of bacteria spp & traits
###5. Feature construction with bacterial traits
###6. Data visualization: summary “state of knowledge” on bacteria causing disease in mammals or humans
###7. Use traits to predict transmissibility and human disease outcomes

###2. Match dx to pathogen spp. names

####Check what fields are in bacteria_species_out2.Rdata
bacteria_species_out2 includes pathogenic and non-pathogenic bacteria
```{r}
#commenting this out for now because it doesn't do anything new
#source("R_graph_bacteria_order_pathogen_status3.R")
```

###read in BacDive data accessed by Wanjia Song

###find out fields with at least threshold % coverage
```{r get_proportion}
load("../DATA/PROCESSED/Data from Backdive-2.RData")

threshold = 0.01
D = merge
dim(D)
D = unique(D)
dim(D)
#rm(merge)
#remove NA values
inds.na = which(!is.na(D$value))
D = D[inds.na,]
id_len=length(unique(D$bacdive_id))
D$new_field = paste(D$section, D$subsection, D$field)
save(D, file = "../DATA/PROCESSED/D.Rdata")
count=as.data.frame(table(D$new_field))#combination of subsection and field
var1=count$Var1
freq=count$Freq
var<-list()
freqs<-list()
for (i in 1:length(freq)) {
     if (freq[i]>=threshold*id_len) {#this will get fields greater than 1%
         var[[i]]=var1[i]
         freqs[[i]]=freq[i]
     }
}

df=do.call(rbind, Map(data.frame, Feature_name=var, Frequency=freqs))

df$Fraction = df$Frequency/id_len
###get those seen at least 20% of time
df20 = subset(df, Fraction >= 0.2)
write.csv(df20, file ="variables_20_percent_coverage.csv")

###get those seen at least 10% of time
df10 = subset(df, Fraction >= 0.1)
write.csv(df10, file ="variables_10_percent_coverage.csv")

###get those seen at least 5% of time
df05 = subset(df, Fraction >= 0.05)
write.csv(df05, file ="variables_5_percent_coverage.csv")

###get those seen at least 1% of time
df01 = subset(df, Fraction >= 0.01)
save(df01, file = "../DATA/PROCESSED/df01.Rdata")
```

#use df01 to subset all data for those variables
```{r subset_01}
load("../DATA/PROCESSED/D.RData")

merge01 = subset(D, new_field %in% df01$Feature_name)
save(merge01, file = "../DATA/PROCESSED/merge01.Rdata")

merge20 = subset(D, new_field %in% df20$Feature_name)
save(merge20, file = "../DATA/PROCESSED/merge20.Rdata")


```

###spread from long to wide
```{r spread}
load("../DATA/PROCESSED/merge01.Rdata")
load("../DATA/PROCESSED/merge20.Rdata")
#merge01 = merge01[c(1:100),]
ptm <- proc.time()

# spreadDf = spread(merge01, key = new_field, value = value)
spreadDf = spread(merge20, key = new_field, value = value)
dim(spreadDf)
print(proc.time() - ptm)

save(spreadDf, file = "../DATA/PROCESSED/spreadDf.Rdata")
```

```{r}
load("../DATA/PROCESSED/merge20.Rdata")

library("dplyr")
library("tidyr")
datasets_20 <- merge20 %>% 
  unlist() %>% 
  bind_rows() %>% #binding many data frames into one
  gather(grouped_category, value, 1:ncol(.)) %>%
  separate(grouped_category, sep = "\\.", 
           into = c("bacdive_id", 
                    "section", #e.g.,morphology_physiology
                    "subsection", #e.g.,cell_morphology
                    "field" #"cell_shape"
                    )) %>%
  
    #Retain only unique/distinct rows from an input tbl.
unique.data.frame()

dim(datasets_20)
```




###SCRATCH below here

###use df10, variables with coverage over 10%, to subset all data
```{r}


# df10 = read.csv("variables_10_percent_coverage.csv")
# D = merge
# dim(D)
# D = unique(D)
# dim(D)
# #rm(merge)
# 
# #id_len=length(unique(D$bacdive_id))
# D$new_field = paste(D$section, D$subsection, D$field)
# count=as.data.frame(table(D$new_field))#combination of 
# 
# merge10 = subset(D, new_field %in% df10$Feature_name)
# save(merge10, file = "merge10.Rdata")

```


##make histogram plot of every variable.
```{r}

# load("merge10.Rdata")
# 
# plot <- ggplot(merge10, aes(x = value))+
#   geom_histogram(stat="count")+
#   theme(axis.text.x = element_text(angle = 90))
# # +
# #   facet_wrap(~new_field)
# 
# pdf('multiple_page_plot_2_per_page.pdf')
# # Plot on multiple pages (output plot to R/Rstudio)
# facet_multiple(plot = plot, 
#                facets = 'new_field', 
#                ncol = 1, 
#                nrow = 2,
#                scales = "free")
# dev.off()
#plot
#ggsave(plot = plot, file = "histogram_10.jpg")
```


##make histogram plot of only quantitative variables of interest
```{r}

# load("merge10.Rdata")
# fieldsCategorized = read.csv("variables_10_percent_coverage - 1.csv")
# fieldsQuantitative = subset(fieldsCategorized,
#                             quantitative.or.categorical == "quantitative")
# 
# dfQuantitative = subset(merge10, new_field %in% fieldsQuantitative$Feature_name)
# 
# #this changes to NA some that have ranges
# dfQuantitative$value=as.numeric(as.character(dfQuantitative$value))
# plot <- ggplot(dfQuantitative, aes(x = value))+
#   geom_histogram()+
#   theme(axis.text.x = element_text(angle = 90))
# # +
# #   facet_wrap(~new_field)
# 
# pdf('multiple_page_plot_quantitative.pdf')
# # Plot on multiple pages (output plot to R/Rstudio)
# facet_multiple(plot = plot, 
#                facets = 'new_field', 
#                ncol = 1, 
#                nrow = 2,
#                scales = "free")
# dev.off()
```


##make histogram plot of only qualitative variables of interest
```{r}

# load("merge10.Rdata")
# fieldsCategorized = read.csv("variables_10_percent_coverage - 1.csv")
# fieldsCategorical = subset(fieldsCategorized,
#                             quantitative.or.categorical == "categorical")
# 
# dfCategorical = subset(merge10, new_field %in% fieldsCategorical$Feature_name)
# 
# plot <- ggplot(dfCategorical, aes(x = value))+
#   geom_histogram(stat = "count")+
#   theme(axis.text.x = element_text(angle = 90))
# # +
# #   facet_wrap(~new_field)
# 
# pdf('multiple_page_plot_categorical.pdf')
# # Plot on multiple pages (output plot to R/Rstudio)
# facet_multiple(plot = plot, 
#                facets = 'new_field', 
#                ncol = 1, 
#                nrow = 2,
#                scales = "free")
# dev.off()
```





####load pathogenic bacteria species. Subset to include only zoonotic
Read in Gideon_Outbreaks_Bacterium_Mammal_Non-Human.csv. Left off: change to 0 empty cells in column Total.Outbreak.Cases..Approximate.
```{r}
source("R_human_pathogen_outbreaks.R")
```



